### 2024-12-17 main.tex
Add Modes class to code and put new derivation into latex file

% section/paper on sampling with sampled priors.
% \documentclass[preprint,linenumbers]{aastex631} # 6.31 fails with \align!!
\documentclass[linenumbers, onecolumn]{aastex63}
\usepackage{xcolor}

\usepackage{natbib}
\usepackage{amsmath}
\usepackage{enumitem}

\shortauthors{Bernstein et al.}

\newcommand{\ie}{\textit{i.e.}}
\newcommand{\eg}{\textit{e.g.}}
\newcommand{\E}{\mathrm{E}}
\newcommand{\eqq}[1]{Equation~(\ref{#1})}
\newcommand\gary[1]{{\color{red} \{\textbf{GMB}: #1\}}}
%\tabletypesize{\footnotesize}
\newcommand{\vecc}{\ensuremath{\mathbf{c}}}
\newcommand{\vecq}{\ensuremath{\mathbf{q}}}
\newcommand{\vecn}{\ensuremath{\mathbf{n}}}
\newcommand{\vecu}{\ensuremath{\mathbf{u}}}
\newcommand{\vecv}{\ensuremath{\mathbf{v}}}
\newcommand{\hatc}{\ensuremath{\hat{\mathbf{c}}}}
\newcommand{\vecP}{\ensuremath{\mathbf{P}}}
\newcommand{\covm}{C}
\newcommand{\matA}{\ensuremath{A}}
\newcommand{\matD}{D}
\newcommand{\matE}{E}
\newcommand{\matF}{F}
\newcommand{\matG}{G}
\newcommand{\matI}{I}
\newcommand{\matT}{T}
\newcommand{\matX}{X}
\newcommand{\matY}{Y}
\newcommand{\matV}{V}
\newcommand{\matLam}{\Lambda}
\newcommand{\proj}{P}  % Projection matrix
\newcommand{\jac}{J}   % Jacobian
\newcommand{\ident}{I}  % identity
\newcommand{\DD}{\Delta_D}
\newcommand{\likeli}{\mathcal{L}}
\newcommand{\trace}{\text{Tr}}
\begin{document}

\title{Sampling Bayesian probabilities given only sampled priors}

\author[0000-0002-8613-8259]{Gary M. Bernstein}
\affiliation{Department of Physics and Astronomy, University of Pennsylvania, Philadelphia, PA 19104, USA}
\email{garyb@upenn.edu}

\author{Troxel,\ldots}

\begin{abstract}
	\vspace{0.2in}
Later.
\end{abstract}
\reportnum{}

\section{Motivation} \label{sec:intro}

Our motivating situation is that we have a vector of observable summary statistics \vecc, such as the binned 2-point correlation functions of cosmic fields, that we are using to constrain a set of parameters of interest \vecq, which are cosmological parameters such as $\Omega_m, \sigma_8,$ etc.  There is a model $\vecc(\vecq,\vecn)$ for the observables which involves the parameters of interest, but also a vector \vecn\ of nuisance parameters, which include the coefficients of some linear expansion of the redshift distributions $n(z)$ of some population of the galaxies being observed:
\begin{equation}
  n(z) = \sum_{k=1}^{M} n_k b_k(z).
  \label{eq:nzbasis}
\end{equation}
The $b_k$ are a set of predetermined basis functions for the redshift distribution.  In our case there are multiple galaxy populations to be characterized, leading to hundreds of parameters $n_k$ to be considered.

We wish to characterized the Bayesian posterior probability
\begin{equation}
  p(\vecq | \vecc) \propto \int dn\, \likeli(\vecc | \vecq, \vecn) p(\vecq) p(\vecn),
\label{eq:posterior}
\end{equation}
where $\likeli(\vecc | \vecq, \vecn)$ is a known likelihood function of the data, and $p(\vecq)$ and $p(\vecn)$ are priors on the parameters.  This posterior is complex enough that it requires approximation by the output of a Markov chain wandering across the space $(\vecq,\vecn).$

The scenario of interest is when \emph{the prior $p(\vecn)$ is itself known only from a set of samples of $\vecn$ from this distribution.} Most MC samplers require that the posterior be an evaluable function of any value of the parameters, and it is the general task of density estimators to convert the samples of $\vecn$ into an evaluable $p(\vecn).$  But when \vecn\ is of high dimension, two problems arise: first, there may be insufficient available samples to create a viable density estimator; second, sampling of the posterior in (\ref{eq:posterior}) becomes infeasible if the MC must traverse a high-dimensional space.

One approach would be to run a new MC chain over \vecq\ for each of the samples we have of \vecn, and then concatenate these to effect marginalization over \vecn.  This is clearly infeasible if a large number of \vecn\ samples are needed to characterize the prior in this space.

For the DES Y3 analyses, \citet{hyperrank} devised a scheme whereby
the samples of \vecn\ are placed in a grid within some $M$-dimensional
unit hypercube $\mathcal{H}$.  The coordinates \vecu\ within the hypercube
are considered the compressed parameters of $n(z),$ and the
decompression function
$\hat{\vecn}(\vecu)$ outputs the $\vecn_\alpha$ sample at the nearest grid point to
  \vecu.  This solves the problem of creating a continuous \vecu\
  domain, and maintains the equal prior probability of each \vecn\
  sample, but the function \emph{output,} and the resultant likelihood
  function of \vecu\, are discontinuous.
Various strategies are proposed to assign the $\vecn_\alpha$ to the grid points in
$\mathcal{H}$ based on summary statistics, to reduce the
discontinuities.  But the function is never smooth.
As a consequence, many MC samplers become quite inefficient in
sampling of the cosmological posterior.  In particular, samplers such as \textsc{MultiNest} that assume continuity are rendered nearly non-functional.  As a result, the Y3 cosmological priors could not be evaluated with this method.  Instead, the \vecn\ samples were not used, and an \textit{ad hoc} $p(\vecn)$ was adopted which allowed only shifts and dilations of the mean $n(z)$ of the \vecn\ samples.

A more rigorous and extremely efficient method of marginalizing over high-dimensional nuisance parameters was proposed by \citet{hans}, for the case where the following restrictions apply:
\begin{enumerate}
\item The likelihood of the observable \vecc\ is normal, $\vecc \sim \mathcal{N}( \hatc, \covm_c),$ with $\covm_c$ fixed.
\item The prior $p(\vecn)$ can also be assumed to be normal, with a mean and covariance matrix $\covm_n$ that in our case could be assigned from the mean and covariance of the samples of \vecn\ we are given.
\item The model \hatc\ can be linearized in \vecn\ about fiducial values $\vecq_0, \vecn_0$ without loss of accuracy exceeding measurement errors, with the derivatives independent of \vecq.
\end{enumerate}
Under these conditions, \citet{hans} show that the marginalization over \vecn\ is equivalent to adding terms to $\covm_c,$ such that any MC process need not sample \vecn\ at all.

We describe here an approach that is algebraically similar to that of \citet{hans}, but does not require 2nd condition of Gaussianity for the nuisance prior.  Our approach is to seek a linear compression of \vecn\ into a lower-dimensional set of parameters \vecu\ that project away variations in \vecn\ that do not influence the likelihood $\likeli.$  Standard density estimators can then be applied to the \vecu\ values implied by the known \vecn\ samples to yield a prior $p(\vecu)$ that can be used for the MC chain of the cosmological posterior.  The model \hatc, and hence $\likeli,$ will be continuous over this low-dimensional \vecu\ space, and marginalization over \vecu\ will yield posterior probabilities very close to marginalization over the original \vecn.

\section{Derivation}\label{sec:deriv}
We assume that we do have a multivariate normal likelihood for the
observables \vecc\, with the mean being some model
$\hatc(\vecq,\vecn)$ and a fixed covariance matrix $\covm_c.$ In this
section we will assume that the \vecn\ vectors have been shifted by
the mean $\vecn_0 \equiv \left\langle \vecn_\alpha \right\rangle$ so
that their mean is zero.
We seek some function $\hat{\vecn}(\vecu)$ of a lower-dimensional
vector \vecu\ which can be substituted for \vecn\ and yield nearly the
same likelihood function for any \vecn\ in the domain spanned by the
samples $\vecn_\alpha,\; \alpha\in 1\ldots N_\alpha,$ that we are
provided.  This means we want a map $\vecn_\alpha\rightarrow
\vecu_\alpha \rightarrow \hat{\vecn}_\alpha$ such that replacing $\vecn$ with
$\hat\vecn$ alters the cosmological inference by much less than the
other uncertainties.  We will implement this by minimizing the
distance in the space $\vecc$ between the
model generated by $\vecn$ and that by $\hat\vecn$, using the
observations' covariance matrix $\covm_c$ as a metric for the
distance.  This is equivalent to the $\chi^2$ of the difference
between the original and compressed models for the data:
\begin{equation} \left\langle \chi^2 \right\rangle
=  \frac{1}{N_\alpha} \sum_\alpha
                                            \left[ \hatc(\vecq,\vecn_\alpha) - \hatc(\vecq,\hat{\vecn}_\alpha) \right]^T
                                            \covm_c^{-1}
                                            \left[ \hatc(\vecq,\vecn_\alpha) - \hatc(\vecq,\hat{\vecn}\alpha) \right].
\label{eq:chisq}
\end{equation}
If the data are in fact drawn from the model $\hatc(\vecq,\vecn)$ with
a Gaussian likelihood, then this is also the mean shift in $-2\log\likeli$
from the compression.  It is \emph{not}, however, equal to the mean
shift of the overall log of the posterior in
\eqq{eq:posterior}---$\left\langle\chi^2\right\rangle$ is serving as a
proxy for this.

We next assume that the compression is linear, $\hat\vecn=\matX\vecn,$
for some matrix that is idempotent ($\matX\matX = \matX$).  If this is
true, then the contribution to \eqq{eq:chisq} at lowest order in
$\vecn-\hat\vecn$ is
\begin{equation}
  \left\langle \chi^2 \right\rangle = \frac{1}{N_\alpha} \sum_\alpha
  \left[ (\matI-\matX)\vecn_\alpha\right]^T \matF  \left[ (\matI-\matX)\vecn_\alpha\right]
  \label{eq:linearized}
\end{equation}
where we use the Jacobian matrix of the model $\hatc$ to define
\begin{equation}
  \matF \equiv
  \left[\frac{\partial\hatc}{\partial\vecn}\right]_{\vecq_0, \vecn_0}
  \covm_c^{-1} \left[\frac{\partial\hatc}{\partial\vecn}\right]_{\vecq_0,
    \vecn_0}^T.
\label{eq:fisher}
\end{equation}
This quantity is also the Fisher matrix over $\vecn$ giving the information
provided by the observations $\vecc.$  In many cases this matrix will be
rank-deficient and/or poorly conditioned, since the observables are not
likely to be very informative on $\vecn$---if they were, we might not be
concerned with establishing a prior on $\vecn$ to begin with.  We will
not, however, require its inverse.

Since $\matX$ is idempotent, we can write
\begin{align}
  \matX & = V_X \proj_M V_X^T, \\
  \matY \equiv \matI-\matX & = V_X \proj_{-M} V_X^T,
\end{align}
where $V_X$ is unitary and the projection matrix $\proj_M$ is defined as
\begin{equation}
  \left(\proj_M\right)_{ij} \equiv
\begin{cases}
                                            1,  &  i=j\le M \\
                                            0,  & \text{otherwise}
\end{cases}
\end{equation}
and $\proj_{-M}=\matI-\proj_M.$  For a chosen rank $M$ of the
transformation matrix $X$, our task becomes to identify the
eigenvectors $V_X$ that minimize
\begin{align}
  \left\langle \chi^2\right\rangle & = \frac{1}{N_\alpha} \sum_\alpha
  \left[ \matY \vecn_\alpha\right]^T \matF  \left[\matY
                                     \vecn_\alpha\right] \\
       & = \trace \left[ \covm_n V_X \proj_{-M} V_X^T \matF V_X
         \proj_{-M} V_X^T \right].
\end{align}
This optimization is easier if we transform first the systematic variables to
$\vecn^\prime =\matT\vecn$ such that $\covm_{n^\prime}=\matI$, \ie\
make the elements of $\vecn$ uncorrelated and unit-variance.  This
is accomplished by finding the eigensystem $\covm_n=\matV_n \matLam_n
\matV_n^T$ and setting $\matT = \matLam_n^{-1/2} \matV_n^T$.  With
this transformation, we are now seeking a different unitary matrix
$\matV_{X^\prime}$ that minimizes
\begin{align}
  \left\langle \chi^2\right\rangle & = \trace \left[ \matI
    V_{X^\prime} \proj_{-M} V_{X^\prime}^T \left[ (T^{-1})^T \matF
      T^{-1} \right] V_{X^\prime} 
    \proj_{-M} V_{X^\prime}^T \right] \\
  & = \trace \left[ \proj_{-M} V_{X^\prime}^T \matV_G \matLam_G \matV_G^T
    V_{X^\prime} \proj_{-M} \right],
  \label{eq:mintr}\\
  \matG \equiv \left(T^{-1}\right)^T \matF   T^{-1} & = \matLam_n^{1/2} \matV_n^T
  \matF \matV_n \matLam_n^{1/2} = \matV_G \matLam_G \matV_G^T.
  \label{eq:defG}
\end{align}
The last line defines a transformed Fisher matrix $\matG$ and its
(non-negative) eigensystem.  Examination of \eqq{eq:mintr} reveals that this quantity
is minimized if the elements of the unitary matrices satisfy
$\matV_{X^\prime}^T \matV_G = \matI \; \Rightarrow \; \matV_{X^\prime} = \matV_G,$ and the eigensystem of $G$ is placed
in order of decreasing eigenvalue $\lambda^G_i.$ The
elements surviving the project functions yield
\begin{equation}
  \left\langle \chi^2\right\rangle = \sum_{i>M} \lambda_i^G.
  \label{eq:chiresid}
\end{equation}
In other words each eigenvalue of the matrix $\matG$ in \eqq{eq:defG}
gives the contribution to $\left\langle\chi^2\right\rangle$ of one
projection (mode) of $\vecn.$

Transforming the solution back into the space of $\vecn$ yields
\begin{align}
  \matX & = \matT^{-1} \matV_G \proj_M \matV_G^T \matT \\
   & = \left[ \matV_n^T \matLam_n^{1/2} \matV_G \proj_M \right] \left[
     \proj_M \matV_G^T \matLam_n^{-1/2} \matV_n^T \right] \\
\label{eq:DE}
        & \equiv \matD \matE.
\end{align}
We thus obtain our optimal encoding/compression using the nonzero rows of matrix
$\matE$ to give
\begin{equation}
  \vecu_\alpha = \matE \vecn_\alpha
  \label{eq:uu}
\end{equation}
and the decoding/reconstruction of the systematic variables as
\begin{equation}
  \hat\vecn_\alpha = \matD \vecu_\alpha.
  \label{eq:UU}
\end{equation}
One can confirm that this procedure yields a compressed representation
$\vecu$ such that $\covm_u = \matI_M,$ the $M$-dimensional
identity matrix.

The previous derivation ignores the possibility that $\covm_n$ is
singular or nearly so, such that taking $\matLam_n^{-1/2}$ in
\eqq{eq:DE} is not possible.  Indeed in our application, it is
\emph{required} that $\covm_n$ be singular, because we have a sum
normalization constraint on the initial $\vecn_\alpha$ values.  Any
such (nearly) zero element $j$ of $\matLam_n$ has a corresponding
eigenvector $\vecv_j$ of the $\vecn$ space which has zero amplitude
in all of the input samples $\vecn_\alpha,$ so that the $\vecn$ are
confined to a subspace---the reconstructed $\hat\vecn$ should also be.
The compressed
representations $\vecu_\alpha$ and reconstructed $\hat\vecn_\alpha$
should therefore be unaffected by the presence of any $\vecv_j$
component.  This can be accomplished in \eqq{eq:DE} by setting element
$j$ of $\matLam_n^{-1/2}$ to zero, as is typically done during
solutions of least-squares problems using singular value
decompositions.

In summary, the procedure for dimensional reduction is:
\begin{enumerate}
  \item Obtain the Fisher matrix $\matF$ of the system defined in
    \eqq{eq:fisher} using derivatives about the mean sample $\vecn_0,$
    plus the covariance matrix $\covm_n$ of the samples.
  \item From the eigensystem $(\matLam_n,\matV_n)$ of $C_n$, form the
    matrix $G$ defined in \eqq{eq:defG} and get its eigensystem
    $(\matLam_G,\matV_G).$  Place the eigenvalues in descending order.
  \item Choose the size $M$ of the compressed representation to be the
    minimum that keeps the $\left\langle\chi^2\right\rangle$ value in
    \eqq{eq:chiresid} below a chosen threshold, presumably $\ll 1.$
  \item The encoding matrix $E$ and decoding matrix $D$ are formed as
    in \eqq{eq:DE}, taking the inverse $\matLam_n^{-1/2}$ to be zero
    for any eigenvalues that are zero (or within roundoff errors).
  \item Compress all incoming (mean-subtracted) samples using \eqq{eq:uu}.  They will
    have unit covariance matrix and zero mean.
  \item Construct a continuous density estimator in $\vecu$ space that
    mimics the finite sample distribution.  If the distribution is
    normal, this becomes the multidimensional unit normal.  There
    is, however,  no \textit{a prior} reason that this must be the case,
    and something like a normalizing flow will be needed to
    approximate this lower-dimensional representation of the prior.
  \item Sample over $\vecu$ space in the Markov Chain that is sampling the
    posterior on the parameters of interest $\vecq,$ using \eqq{eq:UU}
    to transform each sample back into a $\hat\vecn$ vector.
  \end{enumerate}

The ability to accomodate non-Gaussian distributions of the
nuisance-parameter space is the principle
advantage of our method over the single-step covariance-inflation
method of \citet{hans}.
But even if the \vecu\ prior is well approximated as multivariate normal,
there are practical advantages of compressing the nuisance 
variables and retaining them in the cosmological Markov chain rather
than using the covariance-inflation method for analytic
marginalization.  One can, for example, examine the posterior
distributions of \vecu\ and see if they are at the edges of the prior,
which would potentially indicate an inconsistency between the data and
the prior.

\section{Application}\label{sec:app}

xxxxxxxxx xxxxxxxxx xxxxxxxxx xxxxxxxxx xxxxxxxxx xxxxxxxxx 
xxxxxxxxx xxxxxxxxx xxxxxxxxx xxxxxxxxx xxxxxxxxx xxxxxxxxx 
xxxxxxxxx xxxxxxxxx xxxxxxxxx xxxxxxxxx xxxxxxxxx xxxxxxxxx 
xxxxxxxxx xxxxxxxxx xxxxxxxxx xxxxxxxxx xxxxxxxxx xxxxxxxxx

  \begin{acknowledgments}

G.M.B. acknowledges support from NSF grant AST-2205808 and \ldots.

\end{acknowledgments}
\begin{thebibliography}{dummy}
%\bibliography
%\bibliographystyle{aasjournal}
\bibitem[Cordero et al.(2022)]{hyperrank} Cordero, J.~P., Harrison, I., Rollins, R.~P., et al.\ 2022, \mnras, 511, 2170. doi:10.1093/mnras/stac147

\bibitem[Hadzhiyska et al.(2020)]{hans} Hadzhiyska, B., Alonso, D., Nicola, A., et al.\ 2020, \jcap, 2020, 056. doi:10.1088/1475-7516/2020/10/056
\end{thebibliography}

\end{document}