### 2025-05-23 authors.tex
post-CWR updates including author list, aastex7

% Author list file generated with: mkauthlist 1.3.0+30.g918973f.dirty 
% mkauthlist -j aastex7 -sb --orcid author_list.csv

\suppressAffiliations
\correspondingauthor {G.~M.~Bernstein}

\author[0000-0002-8613-8259, gname='Gary', sname='Bernstein']{G.~M.~Bernstein}
\affiliation{Department of Physics and Astronomy, University of Pennsylvania, Philadelphia, PA 19104, USA}
\email{garyb@physics.upenn.edu}

\author[0000-0002-9719-1717, gname='William', sname='Assignies Doumerg']{W.~d'Assignies D.}
\affiliation{Institut de F\'{\i}sica d'Altes Energies (IFAE), The Barcelona Institute of Science and Technology, Campus UAB, 08193 Bellaterra (Barcelona) Spain}
\email{wdoumerg@igfae.es}

\author[0000-0002-5622-5212, gname='Michael A.', sname='Troxel']{M.~A.~Troxel}
\affiliation{Department of Physics, Duke University Durham, NC 27708, USA}
\email{michael.a.troxel@gmail.com}

\author[0000-0001-8505-1269, gname='Alex', sname='Alarcon']{A.~Alarcon}
\affiliation{Institute of Space Sciences (ICE, CSIC),  Campus UAB, Carrer de Can Magrans, s/n,  08193 Barcelona, Spain}
\email{alexalarcongonzalez@gmail.com}

\author[0000-0002-6445-0559, gname='Alexandra', sname='Amon']{A.~Amon}
\affiliation{Department of Astrophysical Sciences, Princeton University, Peyton Hall, Princeton, NJ 08544, USA}
\email{alexandra.amon@princeton.edu}

\author[0000-0002-3730-1750, gname='Giulia', sname='Giannini']{G.~Giannini}
\affiliation{Kavli Institute for Cosmological Physics, University of Chicago, Chicago, IL 60637, USA}
\email{giulia.giannini15@gmail.com}

\author[0000-0001-7005-8820, gname='Boyan', sname='Yin']{B.~Yin}
\affiliation{Department of Physics, Carnegie Mellon University, Pittsburgh, Pennsylvania 15312, USA}
\email{byin@andrew.cmu.edu}

\author[0000-0002-7069-7857, gname=Sahar, sname=Allam]{S.~S.~Allam}
\affiliation{Fermi National Accelerator Laboratory, P. O. Box 500, Batavia, IL 60510, USA}
\email{sallamatfnalgov@gmail.com}

\author[0000-0003-0171-6900, gname='Felipe', sname='Andrade-Oliveira']{F.~Andrade-Oliveira}
\affiliation{Physik-Institut, University of Zürich, Winterthurerstrasse 190, CH-8057 Zürich, Switzerland}
\email{felipeaoli@gmail.com}

\author[0000-0002-8458-5047, gname='David', sname='Brooks']{D.~Brooks}
\affiliation{Department of Physics \& Astronomy, University College London, Gower Street, London, WC1E 6BT, UK}
\email{david.brooks@ucl.ac.uk}

\author[0000-0003-3044-5150, gname='Aurelio', sname='Carnero Rosell']{A.~Carnero~Rosell}
\affiliation{Instituto de Astrofisica de Canarias, E-38205 La Laguna, Tenerife, Spain}
\affiliation{Laborat\'orio Interinstitucional de e-Astronomia - LIneA, Av. Pastor Martin Luther King Jr, 126 Del Castilho, Nova Am\'erica Offices, Torre 3000/sala 817 CEP: 20765-000, Brazil}
\affiliation{Universidad de La Laguna, Dpto. Astrofísica, E-38206 La Laguna, Tenerife, Spain}
\email{aurelio.crosell@gmail.com}

\author[0000-0002-3130-0204, gname='Jorge', sname='Carretero']{J.~Carretero}
\affiliation{Institut de F\'{\i}sica d'Altes Energies (IFAE), The Barcelona Institute of Science and Technology, Campus UAB, 08193 Bellaterra (Barcelona) Spain}
\email{jorgecarreteropalacios@gmail.com}

\author[0000-0002-7731-277X, gname='Luiz', sname='da Costa']{L.~N.~da Costa}
\affiliation{Laborat\'orio Interinstitucional de e-Astronomia - LIneA, Av. Pastor Martin Luther King Jr, 126 Del Castilho, Nova Am\'erica Offices, Torre 3000/sala 817 CEP: 20765-000, Brazil}
\email{ldacosta@fnal.gov}

\author[0000-0000-0000-0000,gname='Maria Elidaiana', sname='da Silva Pereira']{M.~E.~S.~Pereira}
\affiliation{Hamburger Sternwarte, Universit\"{a}t Hamburg, Gojenbergsweg 112, 21029 Hamburg, Germany}
\email{elidaiana.sp@gmail.com}

\author[0000-0001-8318-6813, gname='Juan', sname='De Vicente']{J.~De~Vicente}
\affiliation{Centro de Investigaciones Energ\'eticas, Medioambientales y Tecnol\'ogicas (CIEMAT), Madrid, Spain}
\email{juan.vicente@ciemat.es}

\author[0000-0000-0000-0000,gname='Spencer', sname='Everett']{S.~Everett}
\affiliation{California Institute of Technology, 1200 East California Blvd, MC 249-17, Pasadena, CA 91125, USA}
\email{spencer.w.everett@jpl.nasa.gov}

\author[0000-0003-4079-3263, gname='Josh', sname='Frieman']{J.~Frieman}
\affiliation{Department of Astronomy and Astrophysics, University of Chicago, Chicago, IL 60637, USA}
\affiliation{Fermi National Accelerator Laboratory, P. O. Box 500, Batavia, IL 60510, USA}
\affiliation{Kavli Institute for Cosmological Physics, University of Chicago, Chicago, IL 60637, USA}
\email{frieman@fnal.gov}

\author[0000-0002-9370-8360, gname='Juan', sname='Garcia-Bellido']{J.~Garc\'ia-Bellido}
\affiliation{Instituto de Fisica Teorica UAM/CSIC, Universidad Autonoma de Madrid, 28049 Madrid, Spain}
\email{juan.garciabellido@uam.es}

\author[0000-0003-3270-7644, gname='Daniel', sname='Gruen']{D.~Gruen}
\affiliation{University Observatory, Faculty of Physics, Ludwig-Maximilians-Universit\"at, Scheinerstr. 1, 81679 Munich, Germany}
\email{dgruen@usm.uni-muenchen.de}

\author[0000-0003-2071-9349, gname='Samuel', sname='Hinton']{S.~R.~Hinton}
\affiliation{School of Mathematics and Physics, University of Queensland,  Brisbane, QLD 4072, Australia}
\email{samuelreay@gmail.com}

\author[0000-0002-9369-4157, gname='Devon L.', sname='Hollowood']{D.~L.~Hollowood}
\affiliation{Santa Cruz Institute for Particle Physics, Santa Cruz, CA 95064, USA}
\email{dhollowo@ucsc.edu}

\author[0000-0002-6550-2023, gname='Klaus', sname='Honscheid']{K.~Honscheid}
\affiliation{Center for Cosmology and Astro-Particle Physics, The Ohio State University, Columbus, OH 43210, USA}
\affiliation{Department of Physics, The Ohio State University, Columbus, OH 43210, USA}
\email{kh@fnal.gov}

\author[0000-0001-5160-4486, gname='David', sname='James']{D.~J.~James}
\affiliation{Center for Astrophysics $\vert$ Harvard \& Smithsonian, 60 Garden Street, Cambridge, MA 02138, USA}
\email{djames44@gmail.com}

\author[0000-0002-8289-740X, gname='Sujeong', sname='Lee']{S.~Lee}
\affiliation{Jet Propulsion Laboratory, California Institute of Technology, 4800 Oak Grove Dr., Pasadena, CA 91109, USA}
\email{sujeong.lee@jpl.nasa.gov}

\author[0000-0003-0710-9474, gname='Jennifer', sname='Marshall']{J.~L.~Marshall}
\affiliation{George P. and Cynthia Woods Mitchell Institute for Fundamental Physics and Astronomy, and Department of Physics and Astronomy, Texas A\&M University, College Station, TX 77843,  USA}
\email{jlm076@tamu.edu}

\author[0000-0001-9497-7266, gname='Juan', sname='Mena-Fernández']{J. Mena-Fern{\'a}ndez}
\affiliation{Universit\'e Grenoble Alpes, CNRS, LPSC-IN2P3, 38000 Grenoble, France}
\email{juan.menafernandez@lpsc.in2p3.fr}

\author[0000-0002-6610-4836, gname='Ramon', sname='Miquel']{R.~Miquel}
\affiliation{Instituci\'o Catalana de Recerca i Estudis Avan\c{c}ats, E-08010 Barcelona, Spain}
\affiliation{Institut de F\'{\i}sica d'Altes Energies (IFAE), The Barcelona Institute of Science and Technology, Campus UAB, 08193 Bellaterra (Barcelona) Spain}
\email{rmiquel@ifae.es}

\author[0000-0002-2598-0514, gname='Andrés', sname='Plazas Malagón']{A.~A.~Plazas~Malag\'on}
\affiliation{Kavli Institute for Particle Astrophysics \& Cosmology, P. O. Box 2450, Stanford University, Stanford, CA 94305, USA}
\affiliation{SLAC National Accelerator Laboratory, Menlo Park, CA 94025, USA}
\email{plazasmalagon@gmail.com}

\author[0000-0002-9646-8198, gname='Eusebio', sname='Sanchez']{E.~Sanchez}
\affiliation{Centro de Investigaciones Energ\'eticas, Medioambientales y Tecnol\'ogicas (CIEMAT), Madrid, Spain}
\email{eusebio.sanchez@ciemat.es}

\author[0000-0003-3054-7907, gname='David', sname='Sanchez Cid']{D.~Sanchez Cid}
\affiliation{Centro de Investigaciones Energ\'eticas, Medioambientales y Tecnol\'ogicas (CIEMAT), Madrid, Spain}
\affiliation{Physik-Institut, University of Zürich, Winterthurerstrasse 190, CH-8057 Zürich, Switzerland}
\email{david.sanchez@ciemat.es}

\author[0000-0002-1831-1953, gname='Ignacio', sname='Sevilla']{I.~Sevilla-Noarbe}
\affiliation{Centro de Investigaciones Energ\'eticas, Medioambientales y Tecnol\'ogicas (CIEMAT), Madrid, Spain}
\email{nsevilla@fnal.gov}

\author[0000-0000-0000-0000,gname='Tae-hyeon', sname='Shin']{T.~Shin}
\affiliation{Department of Physics and Astronomy, Stony Brook University, Stony Brook, NY 11794, USA}
\email{babohahaman@gmail.com}

\author[0000-0002-3321-1432, gname='Mathew', sname='Smith']{M.~Smith}
\affiliation{Physics Department, Lancaster University, Lancaster, LA1 4YB, UK}
\email{mat.smith@soton.ac.uk}

\author[0000-0002-7047-9358, gname='Eric', sname='Suchyta']{E.~Suchyta}
\affiliation{Computer Science and Mathematics Division, Oak Ridge National Laboratory, Oak Ridge, TN 37831}
\email{eric.d.suchyta@gmail.com}

\author[0000-0000-0000-0000,gname='Molly', sname='Swanson']{M.~E.~C.~Swanson}
\affiliation{Center for Astrophysical Surveys, National Center for Supercomputing Applications, 1205 West Clark St., Urbana, IL 61801, USA}
\email{molly.swanson@gmail.com}

\author[0000-0001-9382-5199, gname='Noah', sname='Weaverdyck']{N.~Weaverdyck}
\affiliation{Department of Astronomy, University of California, Berkeley,  501 Campbell Hall, Berkeley, CA 94720, USA}
\affiliation{Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley, CA 94720, USA}
\email{nweaverd@umich.edu}

\author[0000-0002-8282-2010, gname='Jochen', sname='Weller']{J.~Weller}
\affiliation{Max Planck Institute for Extraterrestrial Physics, Giessenbachstrasse, 85748 Garching, Germany}
\affiliation{Universit\"ats-Sternwarte, Fakult\"at f\"ur Physik, Ludwig-Maximilians Universit\"at M\"unchen, Scheinerstr. 1, 81679 M\"unchen, Germany}
\email{jochen.weller@lmu.de}

\author[0000-0002-3073-1512, gname='Philip', sname='Wiseman']{P.~Wiseman}
\affiliation{School of Physics and Astronomy, University of Southampton,  Southampton, SO17 1BJ, UK}
\email{P.S.Wiseman@soton.ac.uk}

% Number in brackets is how many author names to put on front page
\collaboration{7}{(DES Collaboration)}